---
title       : Maximum Pain calculator
subtitle    : Data application for fun and profit
author      : Eusebio Rufian-Zilbermann
job         : Investment Enthusiast
framework   : io2012        # {io2012, html5slides, shower, dzslides, ...}
highlighter : highlight.js  # {highlight.js, prettify, highlight}
hitheme     : tomorrow      # 
widgets     : []            # {mathjax, quiz, bootstrap}
mode        : standalone # {selfcontained, draft}
knit        : slidify::knit2slides
---

## What is Maximum Pain theory?
In the stock market, some option traders (typically larger ones) use techniques 
for risk management known as [delta neutral and delta hedging](https://en.wikipedia.org/wiki/Delta_neutral). These operations create a relationship between stock prices and open positions on options. Maximum Pain is the value that maximizes option buyer's losses and the theory predicts that stock prices on expiration day will trend towards this value.

## Calculating Buyers' Gain

Open Interest is the amount of option contracts that have been purchased and sold for specific stock prices (the strike price) and expiration dates. Call contracts give the buyer the right to buy at the strike price. Call buyers make money based on the difference betweeen stock price and strike price when the stock price is above the strike, otherwise they don't make any money. Put contracts give a right to sell and contract buyers make the difference between strike and option price when the price is below the strike price, zero otherwise. For each strike price we can calculate the Call and Put buyers' gains and aggregate them.

--- .class #id 
### Example Gain calculation
```{r, echo=FALSE, warning=FALSE, message=FALSE}
require(quantmod)
```

```{r,cache=TRUE,fig.width=14,fig.height=4}
BuyerGains<-function(Symbol, Expiration){  
  oc<-getOptionChain(Symbol, Exp=Expiration)  
  CallValue<-function(x){sum((oc$calls$Strike<x)*(x-oc$calls$Strike)*oc$calls$OI)/10000} ## Million $
  PutValue<-function(x){sum((oc$puts$Strike>x)*(oc$puts$Strike-x)*oc$puts$OI/10000)} 
  AllStrikes<-sort(unique(c(oc$calls$Strike,oc$puts$Strike)))
  data.frame(puts=sapply(AllStrikes,PutValue),calls=sapply(AllStrikes,CallValue),row.names=AllStrikes)}
buyerGains<-BuyerGains("FB", as.Date("2015-10-9")) ##Date must be adjusted if run after Oct 9th 2015
barplot(t(data.matrix(buyerGains)),col=c("red","green"),main="Buyer Gains",ylab="Millions of US$",
        xlab="Stock price at expiration in US$",legend.text=c("Put buyers' gain","Call buyers' gain"))
```

--- .class #id 
## Example Maximum Pain calculation
Maximum pain will occur at the strike price where the aggregate Call and Put buyers' gain is minimized
```{r}
StrikePrices<-row.names(buyerGains)
MaxPain<-StrikePrices[which.min(buyerGains[,"calls"]+buyerGains[,"puts"])]
MaxPain
```
## Why Maximum Pain?
When speculating on stock prices it is very important to get a trading edge edge: a way of assisting the decision making process so that the expected total gains are greater than the expected total losses.  

Maximum pain is a tool that can help **YOU** in getting that trading edge and make more money.  
  
  
* * *
Additional info: www.maximum-pain.com